---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# surveydataset

<!-- badges: start -->
[![lifecycle](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.7013164.svg)](https://zenodo.org/record/7013164#.YwFmc3ZBzIU)
[![devel-version](https://img.shields.io/badge/devel%20version-0.1.0-blue.svg)](https://github.com/dataobservatory-eu/surveydataset)
[![dataobservatory](https://img.shields.io/badge/ecosystem-dataobservatory.eu-3EA135.svg)](https://dataobservatory.eu/)
[![Follow rOpenGov](https://img.shields.io/twitter/follow/ropengov.svg?style=social)](https://twitter.com/intent/follow?screen_name=ropengov) 
[![Follow author](https://img.shields.io/twitter/follow/digitalmusicobs.svg?style=social)](https://twitter.com/intent/follow?screen_name=digitalmusicobs)
[![Codecov test coverage](https://codecov.io/gh/dataobservatory-eu/surveydataset/branch/master/graph/badge.svg)](https://app.codecov.io/gh/dataobservatory-eu/surveydataset?branch=master)
[![R-CMD-check](https://github.com/dataobservatory-eu/surveydataset/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dataobservatory-eu/surveydataset/actions/workflows/R-CMD-check.yaml)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/dataobservatory-eu/surveydataset?branch=master&svg=true)](https://ci.appveyor.com/project/dataobservatory-eu/surveydataset)
<!-- badges: end -->

The goal of the 'surveydataset' package is to create special datasets for social sciences surveys that are easier to release, exchange and reuse. 

## Installation

You can install the development version of surveydataset like so:

``` r
remotes::install_github('dataobservatory-eu/surveydataset')
```

## Read in a file

```{r example, message=FALSE}
library(surveydataset)
require(haven)
path <- system.file("ZA5688_v6_sample.sav", package = "surveydataset")
survey_raw       <- read_spss_declared(path)
```

We have created a wrapper around [haven::read_spss()](https://haven.tidyverse.org/reference/read_spss.html) and `declared::as_declared()` to read in surveys in a less ambigous format:

```{r declared}
summary(as.factor(survey_raw$d25))
summary(survey_raw$d25)
```
## Remove constans

```{r reduce-survey}
my_survey <- survey_raw[, c(1:5, 7, 10:12, 33)] # Only to keep example easier to read
head(my_survey)
```

```{r remove-constants}
my_survey <- constant_to_attributes(x = my_survey, constant_columns = 1:5)
head(my_survey)
```

```{r dataset-level-attributes}
dataset_attributes <- attr(my_survey, "datasetAttributes")
dataset_attributes
```

```{r}
library(dataset)
my_survey <- dataset(
  my_survey,
  Dimensions = c("isocntry", "d25"),
  Measures = "qb1_1", 
  Attributes = "wextra", "caseid", 
  Identifier =  dataset_attributes[which(dataset_attributes$attribute == "doi"),2], 
  Title = "Eurobarometer 79.2", 
  Publisher = "GESIS", 
  Creator = person("GESIS"))

version(my_survey) <- dataset_attributes[which(dataset_attributes$attribute == "version"),2]

#my_survey <- dataset_title_create(
#Title = c("EUROBAROMETER 79.2
# APRIL-MAY 2013",dataset_attributes[which(dataset_attributes$attribute == "studyno1"),2]), 
#titleType = c("Title", "AlternativeTitle"))

```

```{r summarize}
summary(my_survey)
```


```{r citation}
utils::toBibtex(bibentry_dataset(my_survey))
```

## Further work

The `surveydataset` package is in a very early stage of its development.  For citations please refer to the current development version as:

```{r citepackage}
citation('surveydataset')
```


## Code of Conduct

Please note that the surveydataset project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
